---
title: "Data_tidying"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries:

```{r}
library(dplyr)
library(tidyr)
library(readr)
```

**Remember! Keyboard shortcut for codeblock is: Ctrl + Alt + i**

Load the salmon by region catch data from the KNB data repository (https://knb.ecoinformatics.org/view/df35b.304.2)

```{r}
catch_original <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
head(catch_original)
```
Lets remove the two last columns, we dont need them. 2 ways to do this (one is hashed out below):

```{r}
catch_data <- catch_original %>% 
  #select(Region, Year, Chinook, Coho, Pink, Chum, Sockeye)
  #or
  select(-All, -notesRegCode)

glimpse(catch_data)

```

Notice Chinook column class is a character instead of numeric!  ^^^^

Try to change it to numeric ....

```{r}
catch_clean <- catch_data %>%
  mutate(Chinook = as.numeric(Chinook))
```

Hmmm... now we need to find spurious characters in that Chinook column. Use the which() and the is.na() functions to look for and list rows in the column that are NA

```{r}
i <- which(is.na(catch_clean$Chinook))
i
```
Instead of going back to the original data and fixing this weird value (and "I" instead of a 1), we can change it here and leave the raw data unedited. This is a more reproducible workflow.

```{r}
catch_clean <- catch_data %>%
  mutate(Chinook = if_else(Chinook == "I", "1", Chinook)) %>%
  mutate(Chinook = as.numeric(Chinook))

glimpse(catch_clean)

i <- which(is.na(catch_clean$Chinook))

i
```

Now we can transform/transpose this data from a wide format to a long format
```{r}
catch_long <- catch_clean %>%
  pivot_longer(cols = (-c(Region, Year)), names_to = "Species", values_to = "Catch")
head(catch_clean)
head(catch_long)
```

Now lets make it wide again
```{r}
catch_wide <- catch_long %>%
  pivot_wider(names_from = Region, values_from = Catch)
head(catch_wide)
```

## Split Apply Combine

mean catch by region

```{r}
mean_region <- catch_long %>%
  group_by(Region) %>%
  summarise(catch_mean = mean(Catch))

head(mean_region)
```

```{r}
mean_region_species <- catch_long %>%
  group_by(Region, Species) %>%
  summarise(catch_mean = mean(Catch), n_years =  n())

head(mean_region_species)
```


 Load the region table, which has more information about each region instead of just a region code
 
```{r}
region_defs <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1") %>% 
    select(code, mgmtArea)

head(region_defs)
```

And then join the catch table to the region table using a LEFT join:

```{r}
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))
head(catch_joined)

```

