---
title: "utqiagvik_birds"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries:

```{r}
library(rvest)
library(janitor)
library(dplyr)
library(tidyr)
library(readr)
```
Load the species code and common name data from a website

```{r}
webpage <- read_html("https://www.pwrc.usgs.gov/bbl/manual/speclist.cfm")

tbls <- html_nodes(webpage, "table") %>% 
  html_table(fill = TRUE)

species <- tbls[[1]] %>% 
  clean_names() %>% 
  select(alpha_code, common_name) %>% 
  mutate(alpha_code = tolower(alpha_code))
```

Load the bird data (from https://arcticdata.io/catalog/view/doi:10.18739/A23R0PT35)

```{r}

predator_csv <- read_csv("https://arcticdata.io/metacat/d1/mn/v2/object/urn%3Auuid%3A9ffec04c-7e2d-41dd-9e88-b6c2e8c4375e")

nest_csv <- read_csv("https://arcticdata.io/metacat/d1/mn/v2/object/urn%3Auuid%3A982bd2fc-4edf-4da7-96ef-0d11b853102d")

```


Check out the tables:

```{r}
head(predator_csv)
head(nest_csv)
head(species)
```

Create a function to join the predator table to the species table.

```{r}
species_func <- function(L.table, R.table){
  joined <- left_join(L.table, R.table, by = c("species" = "alpha_code"))
  return(joined)
}
```

Run the function on the predator and species table

```{r}
predator_joined <- species_func(predator_csv, species)
head(predator_joined)
```

Then analyze total number of predators by species and year

```{r}
sum_preds_by_year <- predator_joined %>%
  group_by(year, species) %>% 
  summarize(pred_count = sum(count))
head(sum_preds_by_year)
```

Then apply the function (again!) to add the common names

```{r}
sum_preds_by_year_joined <- species_func(sum_preds_by_year, species)
head(sum_preds_by_year_joined)
```

```{r}
sum_eggs_pred_by_year_species <- nest_csv %>%
  group_by(year, species) %>% 
  summarize(total_predated = sum(number_eggs_predated))
head(sum_eggs_pred_by_year_species)
```
Then apply the function (again!) to add the common names

```{r}
sum_eggs_pred_by_year_species_joined <- species_func(sum_eggs_pred_by_year_species, species)
head(sum_eggs_pred_by_year_species_joined)
```
```{r}

```

